Exploring Bounds

### 2015-04-16 draft.tex
added some plotkin examples

\newtheorem{theorem}{Theorem}
\numberwithin{theorem}{subsection}

@@

The Plotkin Bound is an upper bound that often improves upon the Sphere Packing Bound on $A_q(n,d)$.  It is limited, however, in that it is only valid when $d$ is sufficiently
close to $n$.  Consider that for $q=2$ our restriction is $n<2d$,  for $q=3$ our restriction is $n<\frac{3}{2}d$, for $q=4$ our restriction is $n<\frac{4}{3}d$, etc.  For large $q$, the
number of codes we can consider becomes increasingly limited.

@@

@@

Note that for $x,y \in \mathcal{C}$, d$(x,y)=0$ for $x=y$, and $d \le$ d$(x,y)$ for $x\neq y$.  This implies that
\begin{equation}
M(M-1)d\le S
\end{equation}

This comes directly from equation \ref{plotkin_1}.  We have $M$ possible codewords to consider for $x$ in the outer summation, $M-1$ distinct codewords from $x$ to consider in the inner summation for $y$, and the distance between $x$ and $y$ will always be greater than or equal to $d$.

Next, let $\mathcal{M}$ be the $M \times n$ matrix whose rows are the codewords of $\mathcal{C}$.  For $1 \le i \le n$, let $n_{i,\alpha}$ be the number of times $\alpha \in
 \mathbb{F}_q$ occurs in column $i$ of $\mathcal{M}$.  Note that $\sum_{\alpha \in \mathbb{F}_q} n_{i,\alpha} = M$ for $1\le i \le n$.  This is simply because there are $M$ total
 elements of $\mathbb{F}_q$ in each column, so the sum over all $n_{i,\alpha}$ must necessarily be $M$.  Consequently, we have

\begin{equation}
\begin{split}
S = \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}(M - n_{i,\alpha}) &= M\sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha} - \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}^2 \\
&= nM^2 - \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}^2
\end{split}
\end{equation}

@@

\begin{equation}
S \le nM^2 - \sum_{i=1}^n q^{-1} \left(\sum_{\alpha\in\mathbb{F}_q} 1\cdot n_{i,\alpha} \right )^2 = nM^2 - \frac{nM^2}{q} = nrM^2
\end{equation}

Combining the two, we get FINISH.

@@

>>>>

\newtheorem{theorem}{Theorem}
%\numberwithin{theorem}{subsection}

\theoremstyle{definition}
\newtheorem{exmp}{Example}
\numberwithin{exmp}{subsection}

@@

The Plotkin Bound is an upper bound that often improves upon the Sphere Packing Bound on $A_q(n,d)$.

@@

Before proving this bound, it is important to note that it is only valid when $d$ is sufficiently close to $n$.  For large $q$, this is can be a considerable limiting factor.
Consider that for $q=2$ our restriction is $n<2d$,  for $q=3$ our restriction is $n<\frac{3}{2}d$, for $q=4$ our restriction is $n<\frac{4}{3}d$, etc.

@@

Note that for $x,y \in \mathcal{C}$, d$(x,y)=0$ for $x=y$, and $d \le$ d$(x,y)$ for $x\neq y$.  This implies that
\begin{equation} \label{plotkin_1_5}
M(M-1)d\le S
\end{equation}

This comes directly from equation \ref{plotkin_1}.  We have $M$ possible codewords to consider for $x$ in the outer summation, $M-1$ distinct codewords from $x$ to consider in the inner summation for $y$, 
and the distance between $x$ and $y$ will always be greater than or equal to the minimum distance $d$.

Next, let $\mathcal{M}$ be the $M \times n$ matrix whose rows are the codewords of $\mathcal{C}$.  For $1 \le i \le n$, let $n_{i,\alpha}$ be the number of times $\alpha \in
 \mathbb{F}_q$ occurs in column $i$ of $\mathcal{M}$.  Note that $\sum_{\alpha \in \mathbb{F}_q} n_{i,\alpha} = M$ for $1\le i \le n$.  This is simply because there are $M$ total
 elements of $\mathbb{F}_q$ in each column, so the sum over all $n_{i,\alpha}$ must necessarily be $M$.

I claim that we can rewrite $S$ as a double sum over these $n_{i,\alpha}$ values.  Specifically,

\begin{equation}
S = \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}(M - n_{i,\alpha}).
\end{equation}

Note that $n_{i,\alpha}(M - n_{i,\alpha})$ is multiplying the number of times $\alpha$ occurs in column $i$ by the number of times it doesn't occur in column $i$.  In other words, we're counting the number of times two elements of the same
column are not equivalent, exactly the condition that constitutes our Hamming distance.   By summing over all $\alpha$ and then all columns, we arrive at the same value generated by \ref{plotkin_1}.

Next, simplify:

\begin{equation}
\begin{split}
\sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}(M - n_{i,\alpha}) &= M\sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha} - \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}^2 \\
&= nM^2 - \sum_{i=1}^{n} \sum_{\alpha\in\mathbb{F}_q} n_{i,\alpha}^2
\end{split}
\end{equation}

@@

\begin{equation} \label{plotkin_2}
S \le nM^2 - \sum_{i=1}^n q^{-1} \left(\sum_{\alpha\in\mathbb{F}_q} 1\cdot n_{i,\alpha} \right )^2 = nM^2 - \frac{nM^2}{q} = nrM^2
\end{equation}

Combining \ref{plotkin_1_5} and \ref{plotkin_2}, we get

\begin{equation}
M(M-1)d \le nrM^2
\end{equation}

which simplifies to

\begin{equation}
M \le \left\lfloor \frac{d}{d-rn} \right\rfloor
\end{equation}

@@

It is interesting to see exactly how the Sphere Packing and Plotkin bounds compare for specific values of $n$ and $d$.  For the following examples we will assume $q=2$.    
Also, recall that if $d$ is even, $A_2(n,d) = A_2(n-1,d-1)$.

\begin{exmp}
$(n,d) = (7,4)$.

Applying the Sphere Packing bound, we have $A_2(7,4) = A_2(6,3) \le \frac{2^6}{1 + 6}$, 
giving us $A_2(7,4) \le 9$.

Applying the Plotkin bound, we have $A_2(7,4) \le \frac{4}{4-\frac{1}{2}\cdot7} = 8$.

\end{exmp}

\begin{exmp}
$(n,d) = (9,6)$.

Applying Sphere Packing, we have  $A_2(9,6) = A_2(8,5) \le \frac{2^8}{1+8+28}$, giving us $A_2(n,d) \le 6$.

Applying the Plotkin bound, we have $A_2(9,6) \le \frac{6}{6-\frac{1}{2}\cdot9} = 4$.
\end{exmp}
